---
title: "UHI Intensity Analysis"
author: "Yulin Gao"
date: "2022-10-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

For the UHI: The ideal configuration is LST_urb_all-LST_rur_all for the entire urbanized area (from the US_Urbanized file) and LST_urb_CT_act-LST_rur_all for individual census tracts within the urbanized areas (from the census file)

```{r import data}
library(RMySQL)
server_info = read.table("database_server_info.txt")
mysqlconnection = dbConnect(RMySQL::MySQL(),
                              dbname=server_info[5,],
                              host=server_info[1,],
                              port=as.integer(server_info[2,]),
                              user=server_info[3,],
                              password=server_info[4,])
df = dbReadTable(conn = mysqlconnection, name = "census_uhi_us_urbanized_recalculated")
```

Select UHI related info from the dataframe

```{r uhi}
index = sapply(colnames(df), function(x) grepl("UHI", x, fixed = TRUE))
uhi_df = df[index]
uhi_df["city"] = df["Urban_name"]
head(uhi_df)
```

Add the county column

```{r county}
county_table = dbReadTable(conn = mysqlconnection, name = "uscity_info")
head(county_table)
city_state = strsplit(uhi_df$city, ",", fixed = TRUE)
city = sapply(city_state, "[", 1)
city = strsplit(city, "-|/|\\(")
city = sapply(city, "[", 1)
state = sapply(city_state, "[", 2)
state = strsplit(state, " |-")
state = sapply(state, "[", 2)
uhi_df["ct"] = city
uhi_df["state"] = state
county_vec = rep("null", nrow(uhi_df))
fips_vec = rep(0, nrow(uhi_df))
for (i in 1:nrow(uhi_df)) {
  ct = uhi_df$ct[i]
  state = uhi_df$state[i]
  city_info = county_table[county_table$city == ct & county_table$state_id == state,]
  if (dim(city_info)[1] != 0) {
    county = city_info$county_name[1]
    county_fips = city_info$county_fips[1]
    county_vec[i] = county
    fips_vec[i] = county_fips
  }
}
uhi_df["county"] = county_vec
uhi_df["county_fips"] = fips_vec
head(uhi_df)
```

Generate the city names with strong heat island effects from the uhi table

```{r}
city_uhi = uhi_df[!duplicated(uhi_df[, "county_fips"]),]
urbanized_county = city_uhi[, c("county", "state", "county_fips")]
urbanized_county = urbanized_county[urbanized_county$county_fips != 0,]
write_csv(urbanized_county, "selected_us_cities.csv")
```

Select county-level uhi data
```{r}
city_uhi = uhi_df[!duplicated(uhi_df[, "county_fips"]),]
city_uhi = city_uhi[, c("county", "state", "county_fips", colnames(city_uhi)[7:12])]
colnames(city_uhi)[3] = "County.Code"
city_uhi = city_uhi[city_uhi$County.Code != "35013",]
colnames(city_uhi)[3] = "county_code"
dbWriteTable(mysqlconnection, value = city_uhi, name = "county_uhi", append = TRUE)
dbDisconnect(mysqlconnection)
```

import the death data

```{r}
library(RMySQL)
server_info = read.table("database_server_info.txt")
mysqlconnection = dbConnect(RMySQL::MySQL(),
                              dbname=server_info[5,],
                              host=server_info[1,],
                              port=as.integer(server_info[2,]),
                              user=server_info[3,],
                              password=server_info[4,])
death_df = dbReadTable(conn = mysqlconnection, name = "cdc_elderly_death")
city_uhi = dbReadTable(conn = mysqlconnection, name = "county_uhi")
population_df = dbReadTable(conn = mysqlconnection, name = "county_elderly_population")
```

```{r}
library(dplyr)
death_df$county_code = as.character(death_df$county_code)
death_df$age_group = as.factor(death_df$age_group)
death_all_age_all_time = death_df %>% group_by(county_code) %>% summarise(total_Death = sum(death))

death_age1 = death_df[death_df$age_group == "55-64", ]
death_age1_all_time = death_age1 %>% group_by(county_code) %>% summarise(total_Death = sum(death))

death_age2 = death_df[death_df$age_group == "65-74", ]
death_age2_all_time = death_age2 %>% group_by(county_code) %>% summarise(total_Death = sum(death))

death_age3 = death_df[death_df$age_group == "75-84", ]
death_age3_all_time = death_age3 %>% group_by(county_code) %>% summarise(total_Death = sum(death))

death_age4 = death_df[death_df$age_group == "85+", ]
death_age4_all_time = death_age4 %>% group_by(county_code) %>% summarise(total_Death = sum(death))
```

Create tables with uhi and death
```{r}
uhi_death_all = merge(city_uhi, death_all_age_all_time, by="county_code")
lm.all = lm(total_Death~.-county_code-county-state, data = uhi_death_all)
summary(lm.all)
lm.all = lm(total_Death~UHI_annual_day_city, data = uhi_death_all)
summary(lm.all)
```

import the population data

```{r}
colnames(population_df)[1] = "county_code"
population_allage_df = population_df %>% group_by(county_code) %>% summarise(population = sum(population))
uhi_death_ratio_all = merge(uhi_death_all, population_allage_df, by="county_code")
uhi_death_ratio_all[, "death_ratio"] = uhi_death_ratio_all$total_Death / uhi_death_ratio_all$population
county_death_ratio = uhi_death_ratio_all[, -c(4:9)]
dbWriteTable(conn = mysqlconnection, value = county_death_ratio, name = "county_death_ratio", append = TRUE)
dbDisconnect(mysqlconnection)
```

Perform regression of death rate on uhi data (lm and randomForest)
```{r}
#lm.ratio = lm(death_ratio~.-County.Code-county-state-total_Death-population, data = uhi_death_ratio_all)
lm.ratio = lm(death_ratio~UHI_summer_day_city, data = uhi_death_ratio_all)
summary(lm.ratio)
# cov(uhi_death_ratio_all$death_ratio, uhi_death_ratio_all$UHI_summer_day_city)
library(randomForest)
rf.fit = randomForest(death_ratio~.-county_code-county-state-total_Death-population, data = uhi_death_ratio_all, ntree=1000, keep.forest=FALSE, importance=TRUE)
rf.fit
```

Plot the importance

```{r}
library(ggplot2)
# Get variable importance from the model fit
ImpData <- as.data.frame(importance(rf.fit))
ImpData$Var.Names <- row.names(ImpData)

ggplot(ImpData, aes(x=Var.Names, y=`%IncMSE`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`%IncMSE`), color="skyblue") +
  geom_point(aes(size = IncNodePurity), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
```


Perform transformation and we plus 2 on uhi data
```{r}
library(car)
data_fit = uhi_death_ratio_all[, c(4:9, 12)]
data_fit[, 1:6] = data_fit[,1:6] +2
colnames(data_fit)[1:6] = c("AD", "AN", "SD", "SN", "WD", "WN")
summary(powerTransform(cbind(AD, AN,SD, SN, WD, WN)~1, data_fit))
data_tr = data_fit
data_tr[, 1:6] = bcPower(data_tr[, 1:6], c(0, 0, 0, -2, -2, 1))
lm.tr = lm(death_ratio~., data = data_tr)
summary(lm.tr)
boxCox(lm.tr)
data_tr[, 7] = bcPower(data_tr[, 7], 2)
lm_2tr = lm(death_ratio~., data = data_tr)
summary(lm_2tr)
```

import the yearly uhi data

```{r}
library(RMySQL)
server_info = read.table("database_server_info.txt")
mysqlconnection = dbConnect(RMySQL::MySQL(),
                              dbname=server_info[5,],
                              host=server_info[1,],
                              port=as.integer(server_info[2,]),
                              user=server_info[3,],
                              password=server_info[4,])
uhi_data = dbReadTable(conn = mysqlconnection, name = "ee_cb_uhi")
county_death_ratio = dbReadTable(conn = mysqlconnection, name = "county_death_ratio")
```

```{r}
uhi_avg = uhi_data %>% group_by(county) %>% summarize(su_day = mean(su_daytime), su_night = mean(su_nighttime), wt_day = mean(win_daytime), wt_night = mean(win_nighttime))
death_ratio_df = county_death_ratio[, c("county_code", "total_Death", "death_ratio")]
colnames(death_ratio_df)[1] = "county"
uhi_ratio_df = merge(uhi_avg, death_ratio_df, by="county")
lm.death = lm(total_Death~.-county-death_ratio, data = uhi_ratio_df)
summary(lm.death)
lm.ratio = lm(death_ratio~.-county-total_Death, data = uhi_ratio_df)
summary(lm.ratio)
```

analyze the uhi data in 2010

```{r}
uhi_2010 = uhi_data[uhi_data$yyyymm == "2010", ]
death_2010 = death_df[grepl("2010", death_df$Month.Code), ]
death_2010_all = death_2010 %>% group_by(County.Code) %>% summarize(total_death = sum(Deaths))
ratio_2010 = merge(death_2010_all, population_allage_df, by="County.Code")
colnames(ratio_2010)[1] = "county"
uhi_ratio_2010 = merge(uhi_2010, ratio_2010, by="county")
uhi_ratio_2010[, "death_ratio"] = uhi_ratio_2010$total_death / uhi_ratio_2010$population
data_fit_2010 = uhi_ratio_2010[, c(3,4,5,6,9)]
lm.2010 = lm(death_ratio~., data=data_fit_2010)
summary(lm.2010)
library(randomForest)
rf.fit = randomForest(death_ratio~., data = data_fit_2010, ntree=1000, keep.forest=FALSE, importance=TRUE)
rf.fit
```

Plot the importance

```{r}
library(ggplot2)
# Get variable importance from the model fit
ImpData <- as.data.frame(importance(rf.fit))
ImpData$Var.Names <- row.names(ImpData)

ggplot(ImpData, aes(x=Var.Names, y=`%IncMSE`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`%IncMSE`), color="skyblue") +
  geom_point(aes(size = IncNodePurity), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
```
```





